%-------------------------
% ATS-Friendly Resume in Latex (2-Page Version)
% Author : Shakul Pathak
% Based on: https://github.com/sb2nov/resume
% License : MIT
%------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[pdftex]{hyperref}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins for better space utilization
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.6in}
\addtolength{\textheight}{1.2in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-6pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

%-------------------------
% Custom commands
\newcommand{\resumeItem}[2]{
  \item\small{
    \textbf{#1}{#2 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-6pt}
}

\newcommand{\resumeSubItem}[2]{\resumeItem{#1}{#2}\vspace{-4pt}}

\renewcommand{\labelitemii}{-}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
%%%%%%  CV STARTS HERE  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%----------HEADING-----------------
\begin{center}
  \textbf{\Large Shakul Pathak} \\
  \vspace{2pt}
  Cambridge, MA 02139 | +1 617-685-9569 | shakulp@mit.edu \\
 LinkedIn: \href{https://www.linkedin.com/in/shakul-pathak/}{linkedin.com/in/shakul-pathak/} | Google Scholar:  \href{https://scholar.google.com/citations?user=6gel9QYAAAAJ}{scholar.google.com/citations?user=6gel9QYAAAAJ}
\end{center}
\vspace{-8pt}

\section{Professional Summary}
\vspace{-2pt}
Modeling Engineer/Researcher specializing in physics-based (SPM, P2D, DFN) and data-driven simulation of electrochemical systems (batteries). PhD candidate at MIT with expertise in electrochemical modeling, Porous Electrode Theory, Digital Twin development and Battery Management Systems (BMS). Proven track record translating models into engineering decisions for Shell USA, Toyota Research Institute, and SLB. co-lead author other major journals with 8+ peer-reviewed publications. Knows effective communication across research and business through conference talks, industry consulting, and coursework at MIT Sloan.

\section{Education}
  \resumeSubHeadingListStart
    \resumeSubheading
      {Massachusetts Institute of Technology (MIT)}{Cambridge, MA}
      {PhD in Chemical Engineering, GPA: 5.0/5.0, Minor in Entrepreneurial Finance (MIT Sloan)}{Sep 2021 - Dec 2025}
    \resumeSubheading
      {Massachusetts Institute of Technology (MIT)}{Cambridge, MA}
      {MS in Chemical Engineering Practice (MSCEP), GPA: 5.0/5.0}{Sep 2021 - May 2024}
    \resumeSubheading
      {Indian Institute of Technology (IIT) Kharagpur}{Kharagpur, India}
      {Bachelor and Master of Technology in Chemical Engineering, CGPA: 9.55/10}{Jul 2016 - Jul 2021}
  \resumeSubHeadingListEnd

%-----------EXPERIENCE-----------------
\section{Professional Experience}
  \resumeSubHeadingListStart
    \resumeSubheading
      {Bazant Research Group (MIT)}{Cambridge, MA}
      {Graduate Researcher, PI: Prof. Martin Z. Bazant}{June 2022 - Present}
      \resumeItemListStart
        \resumeItem{Digital Twin Models: }{Developed lean (5x parameter reduction) SPM and P2D battery models for rapid State of Health (SOH) and State of Charge (SOC) tracking. Framework supports Battery Management Systems (BMS) development and predictive maintenance.}
        \resumeItem{Model-guided Discovery of Charge-Transfer Kinetics: }{Led model-guided design of experiments (DOE) across 4 experimental teams for model parameterization of LCO, NMC, and LFP cathodes. Developed Electrochemical Impedance Spectroscopy (EIS) software for learning kinetic models from frequency-domain data. Work sponsored by Shell USA \& Toyota Research Institute (TRI).}
        \resumeItem{Rapid Estimation of Battery Deformation: }{Co-developed Julia Finite Volume Method (FVM) software for millisecond-scale (100x speedup compared to vanilla simulators) battery swelling simulation. Validated against 20Ah NCM-Graphite pouch cell data for battery pack design optimization.}
        \resumeItem{Quantifying Microstructure-Performance Relationships: }{Built multiscale FVM models for phase-change LFP materials. Developed graph framework reducing simulation time from weeks to hours. Created impedance models for polycrystalline NCM cathodes for deconvoluting processes that contribute to low to moderate (< 2C) rate response. Work sponsored by MathWorks Fellowship.}
        \resumeItem{Degradation and Diagnostics: }{Developed interpretable model for dead lithium dynamics limiting Li-metal battery cycle life. Co-developed method combining model features into degradation metrics for SOH estimation (provisional patent filed).}
      \resumeItemListEnd

    \resumeSubheading
      {SLB (Formerly Schlumberger)}{Cambridge, MA}
      {Research Consultant}{Jan 2023 - March 2023}
      \resumeItemListStart
        \resumeItem{Direct Lithium Extraction Analysis: }{Quantified bottlenecks in Direct Lithium Extraction (DLE) systems for unlocking 70\% recovery. Root cause analysis using transport phenomena modeling supported SLB's Nevada project decisions.}
        \resumeItem{Hydrogen SOEC Feasibility: }{Ran physics-guided feasibility assessment of integrating electrolyzer in traditional industries, spotted promising entry points and quantified lifetime energy and \$ savings per installation. Evaluated capex/opex sensitivity and emissions reduction for go/no-go decisions.}
      \resumeItemListEnd

    \resumeSubheading
      {AstraZeneca}{Gaithersburg, MD}
      {Research Consultant}{March 2023 - May 2023}
      \resumeItemListStart
        \resumeItem{Automated Testing System: }{Built automated (Arduino) injector assessment platform to automate standardized testing, developing toward animal cruelty-free, low cost and easy compliance testing systems.}
        \resumeItem{ML Quality Control: }{Built Machine Learning models for bioreactor quality control in monoclonal antibody (mAb) production, delivering hybrid algorithms to reduce waste while maintaining yield.}
      \resumeItemListEnd
  \resumeSubHeadingListEnd

\section{Technical Skills}
  \resumeSubHeadingListStart
    \resumeSubItem{Battery Modeling: }{Single Particle Model (SPM), Pseudo-2D Model (P2D), Doyle-Fuller-Newman (DFN), Porous Electrode Theory, Electrochemical Impedance Spectroscopy (EIS), Digital Twin Models, Physics-Based Modeling, Data-Driven Modeling, Multiscale Modeling, Degradation Modeling, Battery Management Systems (BMS), Finite Volume Method (FVM), SOC/SOH Estimation, Capacity Fade Analysis, Cycle Life Prediction}

    \resumeSubItem{Battery Materials: }{Lithium-ion Batteries, Li-metal Batteries, LCO/NMC/LFP/LMFP Cathodes, Graphite Anodes, Phase-Change Materials, Polycrystalline Cathodes, Solid Electrolyte Interphase (SEI), Battery Pack Design, Thermal Management}

    \resumeSubItem{Software \& Programming: }{Python, MATLAB, Julia, C++, TypeScript, MPET, PETLION, PyBaMM, COMSOL Multiphysics, Simulink, LAMMPS, Aspen Plus, Gurobi, Git, LaTeX, Arduino}

    \resumeSubItem{Data Science: }{Machine Learning (ML), Artificial Intelligence (AI), RAG-LLM, FAISS, Design of Experiments (DOE), Statistical Analysis, Regression Analysis, Monte Carlo Methods, Optimization Algorithms}

    \resumeSubItem{Engineering: }{Transport Phenomena, Thermodynamics, Reaction Engineering, Systems Engineering, Process Optimization, Root Cause Analysis, Feasibility Analysis, Capex/Opex Analysis, Heat Transfer}
  \resumeSubHeadingListEnd

\section{ML/AI Projects}
  \resumeSubHeadingListStart
    \resumeSubItem{LLM for Materials Science (2025): }{Built RAG-LLM app using FAISS for X-Ray Absorption Spectroscopy (XAS) analysis with Shirley Meng's group (University of Chicago). Paper to be published in RSC Digital Discovery.}

    \resumeSubItem{Sundai Hack (2025): }{Built knot.ai voice AI for multilingual job matching at 10x cost reduction. Supported by Kuo Sharper Fellowship.}

    \resumeSubItem{Shell AI Hackathon (2024): }{Built ML pipeline for vehicle fleet carbon footprint optimization (Top 10 public leaderboard).}
  \resumeSubHeadingListEnd

\section{Select Publications}
  \resumeSubHeadingListStart
    \resumeSubItem{}{Pathak, S. (co-lead); Zhang, Y. (co-lead); Fraggedakis, D. (co-lead); Gao, T. (co-lead) et al. "Lithium-ion intercalation by coupled ion-electron transfer", Science (2025), DOI:10.1126/science.adq2541}

    \resumeSubItem{}{Pathak, S. (co-lead); Ipers, G. (co-lead); Jiao, J. (co-lead); et al. "Rapid Simulation of Electro-Chemo-Mechanical Deformation of Li-ion Batteries", Nano Letters (2024), DOI:10.1149/1945-7111/ad4f1e}

    \resumeSubItem{}{Pathak, S.; Bazant, M.; "Electrochemical migration of isolated Li-metal islands", In preparation}

    \resumeSubItem{}{Pathak, S.; Bazant, M.; "Analytical approximation of Porous Electrode Theory for Reaction-limited Batteries", In prep}

    \resumeSubItem{}{Ombrini, P.; Pathak, S.; et al. "Modeling Single-Crystal Electrodes as Network of Primary Particles", Energy Environ. Sci. (2025), DOI:10.1039/D5EE04131G}

    \resumeSubItem{}{Co-author; "Current-dependent Morphologies in Li-O2 Batteries by Coupled Ion-Electron Transfer", ACS Applied Materials \& Interfaces (2025), DOI:10.1021/acsami.5c03369}

    \resumeSubItem{}{Shah, A.; Pathak, S.; et al. "Universal approximation for conductance blockade in nanopore membranes", Nano Letters (2024), DOI:10.1021/acs.nanolett.3c04997}

    \resumeSubItem{}{Pathak, S.; Chakraborty, M.; DasGupta, S. "Molecular Investigation of Electrowetted Nanodroplets", Langmuir (2022), DOI:10.1021/acs.langmuir.1c03037}

    \resumeSubItem{}{8+ total publications. Full list: scholar.google.com/citations?user=6gel9QYAAAAJ}
  \resumeSubHeadingListEnd

\section{Awards and Achievements}
  \resumeSubHeadingListStart
    \resumeSubItem{MathWorks Fellow (2025-26): }{MIT School of Engineering fellowship for LFP battery modeling research}

    \resumeSubItem{Kuo Sharper Fellow (2024-25): }{MIT Sloan fellowship for aspiring entrepreneurs and technology leaders}

    \resumeSubItem{MIT 100K Competition (2024): }{Top 13 finalist pitches at MIT 100K Accelerate 2024}

    \resumeSubItem{Research Fellowships: }{DAAD WISE Scholarship (2019, 2020), Mitacs fellowship with 4000+ CAD support (2019)}

    \resumeSubItem{Competitive Exams: }{Top 1.659\% (155K candidates) JEE Advanced 2016; Rank 37 (127K candidates) WBJEE 2016}

    \resumeSubItem{Press: }{Featured in MIT News (2025), Quantum Zeitgeist (2025), Azom for battery modeling research}
  \resumeSubHeadingListEnd

\section{Coursework}
  \resumeSubHeadingListStart
    \resumeSubItem{Core: }{Transport Phenomena, Thermodynamics, Numerical Methods, Reaction Engineering, Systems Engineering, Electrochemical Energy Systems, Complex Analysis, Continuum Mechanics}

    \resumeSubItem{Business: }{Managerial Finance (DCF, IRR, option pricing), New Enterprises (market strategy, competitive moats), Money for Startups (VC, term sheets)}
  \resumeSubHeadingListEnd
\section{Hobbies}
Hiking, long walks, lofi music, trying new food, mentoring, pure/applied mathematics (blog on my \href{https://oscuro-phoenix.github.io/math-for-fun/}{github}), exploring AI for science/productivity.
\end{document}